---
title: "Homework 1"
author: 'Raphael Morsomme'
date: "05-16-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

## Task 1a

```{r}
is_prime <- function(x) {
  
  # Check that x is a positive integer of length 1
  if (length(x)!=1){
    stop("'x' must be of length 1")
  }
  if (!is.numeric(x)){
    stop("'x' must be a number")
  }  
  if (x != floor(x)){
    stop("'x' must be an integer")
  }
  if (x<=0){
    stop("'x' must be positive")
  }          
  
  # Special cases: 1, 2 and 3
  if (x == 1)         return(FALSE)
  if (x %in% c(2,3))  return(TRUE)
  
  # 'divisible' is a logical vector indicating 
  # if x is divisible by the corresponding integer.
  # It is ufficient to only check integer between 2 and square root of x.
  # The code is vectorized for speed.
  divisible <- x %% 2:sqrt(x) == 0
  if (any(divisible)){
    return(FALSE)
  } else {
    return(TRUE)
  }
}
```

## Task 1b

```{r}
is_anagram <- function(x, y) {
  
  # Verify that x and y are characters of same length
  if (!is.character(x)){
    stop("'x' must be a character")
  }
  if (!is.character(y)){
    stop("'y' must be a character")
  }
  
  x_split <- strsplit(x, "")[[1]]
  y_split <- strsplit(y, "")[[1]]
  if (length(x_split) != length(y_split)){
    stop("'x' and 'y' must have the same length")
  }
  
  # Verify that x and y only contain letters (assume lower case)
  is_letter <- c(x_split, y_split) %in% letters
  if (!all(is_letter)){
    stop("'x' and 'y' must only contain letters")
  }
  
  # Check whether x and y are anagrams
  letter_match <- sort(x_split) == sort(y_split)
  if(all(letter_match)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
  
}
```

## Task 2

#### Valid inputs: `is_prime()`

```{r valid_inputs_1a}
is_prime(x = 2)
is_prime(x = 50)
is_prime(x = 1000)
is_prime(x = 13)
```

```{r valid_inputs_1a_additional}
is_prime(x = 1)   # special case
is_prime(x = 1e14) # fast
is_prime(x = 99999999977) # known large prime number from
# http://compoasso.free.fr/primelistweb/page/prime/liste_online_en.php
is_prime(x = 99999999977+1)
```

#### Invalid inputs: `is_prime()`

```{r invalid_inputs_1a, error=TRUE}
is_prime(x = -2)
is_prime(x = pi)
is_prime(x = c(0, 10, 3))
is_prime(x = "a")
```

```{r invalid_inputs_1a_additional, error=TRUE}
is_prime(x = 1/2)
is_prime(x = "1")
```

#### Valid inputs: `is_anagram()`

```{r valid_inputs_1b}
is_anagram(x = "live", y = "evil")
is_anagram(x = "apple", y = "grape")
is_anagram(x = "a", y = "a")
is_anagram(x = "plum", y = "lump")
```

#### Invalid inputs: `is_anagram()`

```{r invalid_inputs_1b, error=TRUE}
is_anagram(x = "small", y = "big")
is_anagram(x = 1, y = "one")
is_anagram(x = "1", y = "1")
is_anagram(x = "apple", "pal")
```

## Task 3

In my function `is_prime`, I first check that the input `x` has the appropriate form. I verify that `x` has length $1$, is a `numeric`, has the form of an integer, and is positive. I then consider three special cases which I handle individually: $1,2$ and $3$. The function then consider all potential divisors of `x`; note that it is sufficient to restrict ourselves to the integers $\{2,3,\dots,\sqrt{x}\}$. I use vectorized code for speed: the $i$th entry of the $(\sqrt{x}-1)$-dimensional logical vector `divisible` indicates whether `x` is divisible by $(i-1)$.

Note that for some cases, a for-loop that iterates through the potential diviors and exits with an early `return()` command may still be faster; for instance, a large `x` which is even could be identified as not prime after the first iteration of the for-loop. But for general `x`, the vectorized form is still extremely fast, even for numbers as large as a $10^{14}$.
